Functional Pearl \\ Programming with Negative, Fractional, and Imaginary Types

### 2012-02-20 dual.tex
imaginary

\title{Functional Pearl \\ Programming with Negative and Fractional Types} 
\authorinfo{Roshan P. James}
           {Indiana University}
           {rpjames@indiana.edu}
\authorinfo{Amr Sabry}
           {Indiana University}
           {sabry@indiana.edu}
\maketitle

\begin{abstract}
  Every functional programmer knows about sum and product types, $a+b$ and
  $a*b$ respectively. Negative and fractional types, $a-b$ and $a/b$
  respectively, are much less known and their computational interpretation is
  unfamiliar and often complicated. We show that in a programming model in
  which information is preserved (such as the model introduced in our recent
  paper on \emph{Information Effects}), negative and fractional types have
  particularly intuitive and natural computational
  interpretations. Intuitively, values of negative types are values that flow
  ``backwards'' to satisfy demands, and values of fractional types are values
  that represent first-class constraints. The combination of negative and
  fractional types allow a programmer to express a wide range of programming
  idioms, including higher-order functions, delimited continuations,
  speculative computation, logic programming, and more.
\end{abstract}

@@

>>>>

\title{Functional Pearl \\ Programming with Negative, Fractional, and Imaginary Types} 
\authorinfo{Roshan P. James}
           {Indiana University}
           {rpjames@indiana.edu}
\authorinfo{Amr Sabry}
           {Indiana University}
           {sabry@indiana.edu}
\maketitle

\begin{abstract}
  Every functional programmer knows about sum and product types, $a+b$ and
  $a*b$ respectively. Negative, fractional, and imaginary types, $a-b$,
  $a/b$, and $a+ib$ respectively, are much less known and their computational
  interpretation is unfamiliar and often complicated. We show that in a
  programming model in which information is preserved (such as the model
  introduced in our recent paper on \emph{Information Effects}), negative,
  fractional, and imaginary types have particularly intuitive and natural
  computational interpretations. Intuitively, values of negative types are
  values that flow ``backwards'' to satisfy demands, values of fractional
  types are values that represent first-class constraints, and values of
  imaginary types are values that capture the essence of recursive datatype
  declarations. The combination of negative, fractional, and imaginary types
  allow a programmer to express a wide range of programming idioms, including
  higher-order functions, delimited continuations, speculative computation,
  logic programming, and more.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Imaginary Types: Intuition}

Imaginary types have also appeared in the literature: we relegate the
connections to Sec.~\ref{related} and proceed with a simple explanation. We
have so far extended the set of types to be the rational numbers. Now we will
push this and extend the set of types to algebraic numbers. In other words,
we will allow datatypes defined by arbitrary polynomials and allow the roots
of such polynomials to be types. Consider the type of binary trees:
\begin{verbatim}
data B = 1 + B * B
\end{verbatim}
Rearranging the terms, this declaration introduces the polynomial $b^2 - b +
1 = 0$ whose roots are $(1 \pm i\sqrt{3}) / 2$.
